---
title: "Economic Sentiment: Error analysis"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
library(here)
library(tidyverse)
```

Read performance data

```{r}
source(here("src/lib/functions.R"))
headlines = read_csv(here("data/raw/gold_sentences.csv")) %>% select(id, headline)
headlines_eng = read_csv(here("data/raw/gold_sentences.csv")) %>% select(id, deepl)
gold = read_csv(here("data/intermediate/gold.csv")) %>% rename(gold=value)
scores = read_csv(here("data/intermediate/combined_predictions.csv"), col_types = "iccccddd")
scores = scores %>% filter(is.na(language) | language != "google", variable != "recessie", variable != "boukes")  %>% mutate(value=trichotomize(value))
names = read_csv(here("src/analysis/report_names.csv"))
sections = setNames(nm=unique(names$section))
variables = setNames(nm=unique(names$variable))
perf = inner_join(scores, gold) %>% inner_join(headlines)
```

# Error analysis of CNN / Deep learning

```{r}
cperf = perf %>% filter(variable=="cnn") %>% 
  group_by(id, gold, headline) %>% summarize(acc=mean(value==gold), value=mean(value), conf=mean(confidence)) 
```

Sentences misclassified as positive

```{r}
cperf %>% filter(gold!=1, value >= 0.8) %>% arrange(-conf)%>% knitr::kable()
```

Sentences misclassified as negative

```{r}
cperf %>% filter(gold!=-1, value <= -0.8) %>% arrange(-conf)%>% knitr::kable()
```

Sentences misclassified as neutral (high confidence)

```{r}
cperf %>% filter(gold!=0, abs(value) <= .2) %>% arrange(-conf) %>% head(10) %>% knitr::kable()
```


Sentences misclassified as neutral (low confidence)

```{r}
cperf %>% filter(gold!=0, abs(value) <= .2) %>% arrange(conf) %>% head(10) %>% knitr::kable()
```

To check embeddings vectors (run in python)

```
import gensim
embeddings = gensim.models.Word2Vec.load("data/tmp/w2v_320d")

def isin(w):
    try:
        embeddings.wv[w]
        return True
    except KeyError:
        return False

isin("dubbelfout")
embeddings.wv.similar_by_word("dubbelfout")
```

# Error analysis of NRC dictionary

Get dictionary, use corpustools to apply at token level

```{r}
nrc <- read_csv(here("data/raw/dictionaries/NRC-Emotion-Lexicon-v0.92-In105Languages-Nov2017Translations.csv")) %>%
  select(c(Eng = `English (en)`, NL = `Dutch (nl)`, Positive, Negative, Fear, Trust))
d = quanteda::dictionary(list(positive = c(nrc$NL[nrc$Positive==1], nrc$NL[nrc$Trust==1]),
                    negative = c(nrc$NL[nrc$Negative==1], nrc$NL[nrc$Fear==1])))

df <- read_csv(here("data/raw/gold_sentences.csv"))
tc = corpustools::create_tcorpus(df %>% select(doc_id=id, text=dutch_lemmas))
tc$code_dictionary(d)
```

Texts that were misclassified as positive 

```{r}
fp = perf %>% filter(variable == "nrc_nl", value ==1,  gold==-1) %>% pull(id)
corpustools::browse_texts(tc$subset(copy=T, subset_meta=doc_id %in% fp), category="code") 
```

Texts that were misclassified as negative

```{r}
fn = perf %>% filter(variable == "nrc_nl", value ==-1,  gold==1) %>% pull(id)
corpustools::browse_texts(tc$subset(copy=T, subset_meta=doc_id %in% fn), category="code") 
```

List of mismatched tokens and sentences in gold standard: misclassified as positive

```{r}
matched_tokens = tc$tokens %>% filter(!is.na(code)) %>% group_by(doc_id, token, code) %>% summarize(n=n()) %>% 
  mutate(id=as.numeric(as.character(doc_id))) %>% ungroup() %>% select(-doc_id)

perf %>% filter(variable == "nrc_nl", value ==1,  gold==-1) %>% select(id, gold, value, headline) %>% 
  inner_join(matched_tokens) %>% group_by(code, token) %>% mutate(n=sum(n)) %>% arrange(-n) %>% filter(code=="positive") %>% 
  knitr::kable()
```

List of mismatched tokens and sentences in gold standard: misclassified as negative

```{r}
perf %>% filter(variable == "nrc_nl", value ==-1,  gold==1) %>% select(id, gold, value, headline) %>% 
  inner_join(matched_tokens) %>% group_by(code, token) %>% mutate(n=sum(n)) %>% arrange(-n) %>% filter(code=="negative")  %>% 
  knitr::kable()
```

# Error analysis of English NRC dictionary

```{r}
de = quanteda::dictionary(list(positive = c(nrc$Eng[nrc$Positive==1], nrc$Eng[nrc$Trust==1]),
                    negative = c(nrc$Eng[nrc$Negative==1], nrc$Eng[nrc$Fear==1])))

tce = corpustools::create_tcorpus(df %>% select(doc_id=id, text=deepl_lemmas))
tce$code_dictionary(de)
```


Texts that were misclassified as positive 

```{r}
fp = perf %>% filter(variable == "nrc_deepl", value ==1,  gold==-1) %>% pull(id)
corpustools::browse_texts(tce$subset(copy=T, subset_meta=doc_id %in% fp), category="code") 
```

Texts that were misclassified as negative

```{r}
fn = perf %>% filter(variable == "nrc_deepl", value ==-1,  gold==1) %>% pull(id)
corpustools::browse_texts(tce$subset(copy=T, subset_meta=doc_id %in% fn), category="code") 
```

List of mismatched tokens and sentences in gold standard: misclassified as positive

```{r}
matched_tokens = tce$tokens %>% filter(!is.na(code)) %>% group_by(doc_id, token, code) %>% summarize(n=n()) %>% 
  mutate(id=as.numeric(as.character(doc_id))) %>% ungroup() %>% select(-doc_id)

perf %>% filter(variable == "nrc_deepl", value ==1,  gold==-1) %>% select(id, gold, value, headline) %>% 
  inner_join(matched_tokens) %>% group_by(code, token) %>% mutate(n=sum(n)) %>% arrange(-n) %>% filter(code=="positive") %>% 
  inner_join(headlines_eng) %>% 
  knitr::kable()
```

List of mismatched tokens and sentences in gold standard: misclassified as negative

```{r}
perf %>% filter(variable == "nrc_deepl", value ==-1,  gold==1) %>% select(id, gold, value, headline) %>% 
  inner_join(matched_tokens) %>% group_by(code, token) %>% mutate(n=sum(n)) %>% arrange(-n) %>% filter(code=="negative")  %>% 
    inner_join(headlines_eng) %>% 
  knitr::kable()
```

List of sentences misclassified by translated that were correct in original

```{r}
nl_correct = perf %>% filter(variable == "nrc_nl", value == gold)
trans_error = perf %>% filter(variable == "nrc_deepl", value != gold) %>% semi_join(nl_correct, by="id") %>% select(id, gold, value_deepl=value, headline) %>% inner_join(headlines_eng)
knitr::kable(trans_error)
```

Browser

```{r}
tce$meta
tce$set_meta("gold_value", gold$gold[match(tce$get_meta("doc_id"), as.character(gold$id))])
tce$set_meta("nrc_deepl", trans_error$value_deepl[match(tce$get_meta("doc_id"), as.character(trans_error$id))])
corpustools::browse_texts(tce$subset(copy=T, subset_meta=doc_id %in% trans_error$id), category="code") 
```

Tokens

```{r}
trans_error %>% 
  inner_join(matched_tokens) %>% group_by(code, token) %>% mutate(n=sum(n)) %>% arrange(-n) %>% filter(code=="positive") %>% 
  knitr::kable()
```

# Naive Bayes features and error analysis

```{r}
f = read_csv(here("data/intermediate/nb_features.csv")) %>% 
  pivot_longer(-feature, names_to="class") %>% mutate(p=exp(value))
intercept = f %>% filter(feature == "___INTERCEPT___") %>% select(class, p)
nbfeatures = f %>% filter(feature != "___INTERCEPT___") %>% inner_join(intercept, by="class", suffix=c("", "_cls")) %>% 
  mutate(pp=p*p_cls) %>% group_by(feature) %>% mutate(max=max(pp), sum=sum(pp)) %>% 
  filter(pp==max) %>% mutate(norm=pp/sum) %>% arrange(-norm) %>% 
  select(feature, class, norm)
nbfeatures %>% head(10) %>% knitr::kable()
```

Positive tokens in texts misclassified as positive

```{r}
matched_tokens = tc$tokens %>% inner_join(nbfeatures, by=c(token="feature")) %>% select(doc_id, token, class, norm) %>% 
   mutate(id=as.numeric(as.character(doc_id))) %>% ungroup() %>% select(-doc_id)
perf %>% filter(variable == "nb", value ==1,  gold==-1) %>% select(id, gold, value, headline) %>% 
  inner_join(matched_tokens) %>% group_by(class, token) %>% arrange(-norm) %>% filter(class==1) %>% 
  knitr::kable()
```

Negative tokens in texts misclassified as negative

```{r}
perf %>% filter(variable == "nb", value ==-1,  gold==1) %>% select(id, gold, value, headline) %>% 
  inner_join(matched_tokens) %>% group_by(class, token) %>% arrange(-norm) %>% filter(class==-1) %>% 
  knitr::kable()
```


Non-neutral tokens in misclassified neutral texts

```{r}
perf %>% filter(variable == "nb", value != 0,  gold==0) %>% select(id, gold, value, headline) %>% 
  inner_join(matched_tokens) %>% group_by(class, token) %>% arrange(-norm) %>% filter(class!=0) %>% 
  knitr::kable()
```
